### 2020-09-30 paper/natastron/methods.tex
Finished Methods

Todo: water down the contents of the methods here.

From our asteroseismic analysis, we report inclination angle, rotational splitting, and rotation period in \textbf{Extended Data} Table 1. The summary statistics on these parameters were taken as the median of the posterior distribution, with uncertainties being the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles. For inclination angle ($i = \arccos(\cos(i))$) and rotation ($P = 1/\nu_{\rm s}$, the full posterior samples were transformed before taking the summary statistics, as our model sampled in $\cos(i)$ and $\nu_{\rm s}$.

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $P_{\rm s}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $P_{\rm s}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications, at the expense of the ability to marginalise for a single value of $P_{\rm s}$ directly.

\subsubsection{Fitting procedure}
The parameter space of the stellar models populations were reduced before calculating the KDEs. These cuts were made in $M$, $\teff$, $\ln(t)$ and \feh, removing any stars in the models that fell more than $3 \times \sigma_{\mathcal{D}}$ outside the observations. Our observables $M$, $t$ and $P$ have asymmetric uncertainties from the Bayesian asteroseismic analysis. We used the larger of the reported uncertainties on each parameter as $\sigma_{\mathcal{D}}$ in each parameter space.

>>>>

In order to extract signatures of stellar rotation from the asteroseismic mode frequencies, we built a model that simultaneously treats the convective background, modes of oscillation, and white noise, while accounting for rotational splitting. The foundations of this approach follows best-practices in asteroseismology \cite{davies+2015}, fitting Harvey profiles for the background \cite{harvey1985}, and expressing the frequencies of the modes using an asymptotic expression \cite{tassoul1980, vrard+2016}.

The first core improvement in this work come in the form of the use of hierarchical models, which account for small-scale deviations in mode frequencies due to effects not explicitly accounted for in our model (such as acoustic glitches \cite{mazumdar+2014}). By improving the inference of mode frequencies in this way, we also improve the ability to resolve the rotational splitting. The second improvement comes in the form of priors on our parameters, and in particular the rotational inclination, which more accurately reflects the true distribution of angles than previous techniques \cite{chaplin+basu2017}. A step-by-step breakdown of the model and all priors can be found in \textbf{Supplementary Materials}.\\

We separated our model fitting into two parts. First, we fit a model for the convective background and white noise only, to the region of the power spectra that did not contain modes of oscillation. This was done using \texttt{PyStan} \cite{vanhoey+2013,carpenter+2017}, for 10,000 iterations on each star\footnote{These \texttt{PyStan} runs were initiated with a random seed of 11, as were all other random processes in this work}. Second, we fit our full model (including the convective background) to the region containing the modes of oscillation. The results of the first fit to the convective background were used as extremely informative priors on the background parameters in this second fit. This was done using \texttt{PyMC3} \cite{vanderwalt+2011,salvatier+2016, thetheanodevelopmentteam+2016} for 2500 iterations each on 4 chains.

From this asteroseismic analysis, we report inclination angle ($i$), rotational splitting ($\nu_{\rm s}$), and rotation period ($P$) in \textbf{Extended Data} Table 1. The summary statistics on these parameters were taken as the median of the posterior distribution, with uncertainties being the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles. For inclination angle ($i = \arccos(\cos(i))$) and rotation ($P = 1/\nu_{\rm s}$), the full posterior samples were transformed before taking the summary statistics, as our model sampled in $\cos(i)$ and $\nu_{\rm s}$.

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $P_{\rm s}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $P_{\rm s}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications, at the expense of the ability to marginalise for a single value of $P_{\rm s}$ directly.\\

The parameter space of the stellar models populations were reduced before calculating the KDEs. These cuts were made in $M$, $\teff$, $\ln(t)$ and \feh, removing any stars in the models that fell more than $3 \times \sigma_{\mathcal{D}}$ outside the observations. Our observables $M$, $t$ and $P$ have asymmetric uncertainties from the Bayesian asteroseismic analysis. We used the larger of the reported uncertainties on each parameter as $\sigma_{\mathcal{D}}$ in each parameter space.
